Precise predictions for same-sign W-bosons scattering at the LHC

### 2018-02-03 Draft/contributions.tex

At tree level, there are three contributions to the $W^+W^+$ production in association with two jets: the pure EW component $\mathcal{O}(\alpha^6)$, the interference $\mathcal{O}(\alphas\alpha^5)$ and the QCD background $\mathcal{O}(\alphas^2\alpha^4)$.
In the present section, the cross sections and distributions are obtained without applying the VBS cuts on $m_{\Pj\Pj}$ and $|\Delta y_{\Pj\Pj}|$.
In Tab.~\ref{tab:LOscanXsec}, the cross sections of the three contributions are displayed.
The EW, QCD, and interference contributions amount to $57\%$, $37\%$, and $6\%$ of the total inclusive cross section, respectively.
The QCD contribution does not posses external gluons due to charge conservation.
Thus the $\mathcal{O}(\alphas^2\alpha^4)$ diagrams only involve gluon exchange in the $t/u$-channel between the quark lines.
This results in a small contribution although the VBS cuts have not been imposed.
The interference between EW and QCD contributions is small, due to color suppression, but not negligible ($t/u$ interference with identical fermions).

@@

In Fig.~\ref{fig:mjjdyjj_1d} these three contributions are shown separately and summed in the differential distribution of the di-jet invariant mass $m_{\Pj\Pj}$ and the rapidity difference $|\Delta y_{\Pj\Pj}|$.
The EW contribution peaks around an invariant mass of about $80\GeV$.
These are due to diagrams where the two jets originate from the decay of a W boson (see middle and right diagrams in Fig.~\ref{diag:LO}).
Note that these contributions are not present in calculation relying on the VBS approximation.
Also, the EW contribution becomes dominant for di-jet invariant mass larger than $500\GeV$ and for jet rapidity difference larger than $2.5$.

@@

This can also be seen in Fig.~\ref{fig:mjjdyjj_2d_LO} where the three contributions are displayed as a function of the di-jet invariant mass and jet rapidity difference.
Again, it is obvious that the region with low di-jet invariant mass and low jet-rapidity difference should be avoided.
This motivates in particular the choice of $m_{\Pj\Pj} > 200\GeV$ and $|\Delta y_{\Pj\Pj}| > 2$ for our inclusive study.
Finally, let us notice that the choice $m_{\Pj\Pj} > 500\GeV$ and $|\Delta y_{\Pj\Pj}| > 2.5$ made by the experimental collaborations is well motivated in order to enhance the EW contribution over its irreducible backgrounds.
This is particularly explicit in Sec.~\ref{subsec:LOfiducial}.

>>>>

At tree level, there are three contributions to the $\PW^+\PW^+$ production in association with two jets: the pure EW component $\mathcal{O}(\alpha^6)$, the interference $\mathcal{O}(\alphas\alpha^5)$, and the QCD background $\mathcal{O}(\alphas^2\alpha^4)$.
In the present section, the cross sections and distributions are obtained without applying the VBS cuts on $m_{\Pj\Pj}$ and $|\Delta y_{\Pj\Pj}|$.
In Tab.~\ref{tab:LOscanXsec}, the cross sections of the three contributions are displayed.
The EW, QCD, and interference contributions amount to $57\%$, $37\%$, and $6\%$ of the total inclusive cross section, respectively.
The QCD contribution does not posses external gluons due to charge conservation.
Thus the $\mathcal{O}(\alphas^2\alpha^4)$ diagrams only involve gluon exchange in the $t/u$-channel between the quark lines.
This results in a small contribution although the VBS cuts have not been imposed.
The interference between EW and QCD contributions is small, due to color suppression, but not negligible ($t/u$ interference with identical fermions).

@@

In Fig.~\ref{fig:mjjdyjj_1d} these three contributions are shown separately and summed in the differential distribution of the di-jet invariant mass $m_{\Pj\Pj}$ and the rapidity difference $|\Delta y_{\Pj\Pj}|$.
The EW contribution peaks around an invariant mass of about $80\GeV$.
These are due to diagrams where the two jets originate from the decay of a W boson (see middle and right diagrams in Fig.~\ref{diag:LO}).
Note that these contributions are not present in calculations relying on the VBS approximation.
Also, the EW contribution becomes dominant for di-jet invariant mass larger than $500\GeV$ and for jet rapidity difference larger than $2.5$.

@@

This can also be seen in Fig.~\ref{fig:mjjdyjj_2d_LO} where the three contributions are displayed as a function of the di-jet invariant mass and jet rapidity difference.
Again, it is obvious that the region with low di-jet invariant mass and low jet-rapidity difference should be avoided.
This motivates in particular the choice of $m_{\Pj\Pj} > 200\GeV$ and $|\Delta y_{\Pj\Pj}| > 2$ for our inclusive study (see below).
Finally, let us notice that the choice $m_{\Pj\Pj} > 500\GeV$ and $|\Delta y_{\Pj\Pj}| > 2.5$ made by the experimental collaborations is well motivated in order to enhance the EW contribution over its irreducible backgrounds.
This is are the cuts used in Sec.~\ref{subsec:LOfiducial}.

### 2018-02-03 Draft/crosssectionLO.tex

\begin{table}[h!]
    \centering
    \begin{tabular}{c|c}
        Code  &  $\sigma[\rm{fb}]$  \\
        \hline
        \hline
        {\sc Bonsay}  &  $X \pm 0.0002$ \\
        {\sc MG5\_aMC}&  $X \pm 0.001$  \\
        {\sc MoCaNLO+Recola}  &  $1.4347 \pm 0.0001$ \\
        {\sc PHANTOM} &  $1.4374 \pm 0.0006 $  \\
        {\sc POWHEG}  &  $1.44092 \pm 0.00009$ \\
        {\sc VBFNLO}  &  $1.43796 \pm 0.00005$ \\
        {\sc Whizard} &  $1.4381 \pm 0.0002 $
    \end{tabular}
    \caption{\label{tab:wg1_LOrates} Rates at LO accuracy within VBS cuts obtained with the different codes used in this comparison,
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process.}
\end{table}

>>>>

\begin{table}[h!]
    \centering
    \begin{tabular}{c|c}
        Code  &  $\sigma[\rm{fb}]$  \\
        \hline
        \hline
        {\sc Bonsay}  &  $X \pm 0.0002$ \\
        {\sc MG5\_aMC}&  $X \pm 0.001$  \\
        {\sc MoCaNLO+Recola}  &  $1.4347 \pm 0.0001$ \\
        {\sc PHANTOM} &  $1.4374 \pm 0.0006 $  \\
        {\sc POWHEG}  &  $1.44092 \pm 0.00009$ \\
        {\sc VBFNLO}  &  $1.43796 \pm 0.00005$ \\
        {\sc Whizard} &  $1.4381 \pm 0.0002 $
    \end{tabular}
    \caption{\label{tab:wg1_LOrates} Rates at LO accuracy within VBS cuts obtained with the different codes used in this comparison,
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process.
    \MP{Please add or check your respective numbers}}
\end{table}

We note that several full predictions are not in statistical agreement.
These are possibly due to too aggressive estimations of statistical uncertainty or ... \MP{To be added}
Nonetheless, all these predictions agree within less than $0.5\%$.
At the level of the cross section, it seems difficult to infer the quality of the the various approximations.
This simply means that the various VBS approximations are not worth than $0.5\%$ at the level of the fiducial cross section at LO for a typical phase-space volume used by experimental collaborations.

### 2018-02-03 Draft/crosssectionNLO.tex
Work on the draft, mainly improvements and corrections

Cross secitons at NLO

\begin{table}[h!]
    \centering
    \begin{tabular}{c|c|c|c}
        Code  &  $\sigma[\rm{fb}]$  \\
        \hline
        \hline
        {\sc Bonsay}  &  $X \pm 0.0009$  \\
        {\sc MG5\_aMC}&  $X  \pm 0.003$  \\
        {\sc MoCaNLO+Recola}  &  $ 1.382 \pm 0.002$ \\
        {\sc POWHEG}  &  $1.3556 \pm 0.0009$  \\
        {\sc VBFNLO}  &  $1.3916 \pm 0.0001$  \\
    \end{tabular}
    \caption{\label{tab:wg1_NLOrates} Rates at NLO-QCD accuracy within VBS cuts obtained with the different codes used in this comparison, 
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process.}
\end{table}
At NLO, rates show slightly larger discrepancies, as it can be observed in Tab.~\ref{tab:wg1_NLOrates}. This is most likely due to low dijet invariant-mass configurations, where
$s-$channel diagrams and interferences are less suppressed than at LO, because of the presence of extra QCD radiation.

>>>>

Cross sections at NLO

\begin{table}[h!]
    \centering
    \begin{tabular}{c|c|c|c}
        Code  &  $\sigma[\rm{fb}]$  \\
        \hline
        \hline
        {\sc Bonsay}  &  $X \pm 0.0009$  \\
        {\sc MG5\_aMC}&  $X  \pm 0.003$  \\
        {\sc MoCaNLO+Recola}  &  $ 1.382 \pm 0.002$ \\
        {\sc POWHEG}  &  $1.3556 \pm 0.0009$  \\
        {\sc VBFNLO}  &  $1.3916 \pm 0.0001$  \\
    \end{tabular}
    \caption{\label{tab:wg1_NLOrates} Rates at NLO-QCD accuracy within VBS cuts obtained with the different codes used in this comparison, 
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process.
    \MP{Please add or check your respective numbers}}
\end{table}

At NLO, the rates show slightly larger discrepancies, as it can be observed in Tab.~\ref{tab:wg1_NLOrates}.
This is most likely due to low di-jet invariant-mass configurations, where $s$-channel diagrams and interferences are less suppressed than at LO, because of the presence of extra QCD radiation.

### 2018-02-03 Draft/distributionsLO.tex

In Fig.~\ref{fig:wg1_mjj-llLO}, we show the distributions in the invariant mass of the tagging-jet (left) and lepton-pair (right).
In both cases we show the absolute distributions in the upper plot, while the lower plot displays the ratio over {\sc VBFNLO}.
For both observables we find a relatively good agreement among the various tools, which confirms the fact that contributions from $s-$channel diagrams as well as from non-resonant configurations are strongly suppressed in the fiducial region.
We have checked that the same level of agreement holds for many other differential distributions.

>>>>

In Fig.~\ref{fig:wg1_mjj-llLO}, we show the distributions in the invariant mass of the tagging-jet (left) and lepton-pair (right).
In both cases we show the absolute distributions in the upper plot, while the lower plot displays the ratio over {\sc VBFNLO}.
For both observables we find a relatively good agreement among the various tools, which confirms the fact that contributions from $s$-channel diagrams as well as from non-resonant configurations are strongly suppressed in the fiducial region.
We have checked that the same level of agreement holds for many other differential distributions.

### 2018-02-03 Draft/scanLO.tex
typocorr
Work on the draft, mainly improvements and corrections

% \subsection{Validity of the VBS approximation}\label{subsec:VBSapprox}

>>>>

As explained previously, the low di-jet invariant mass and low jet rapidity separation is dominated by tri-bosons production.
Therefore a comparison of the various approximation in an \emph{inclusive} phase-space volume.
To that end, we have chosen for the inclusive fiducial volume to take the following values for the VBS cuts

\begin{equation}
\label{eq:inclusive}
	m_{jj} > 200\GeV\,,\qquad |\Delta y_{jj}| > 2 .
\end{equation}

The figure is Fig.~\ref{fig:ratio2d_LO}, the ratio of double-differential cross sections in the plan $\left(m_{\Pj\Pj}, \Delta y_{\Pj\Pj}\right)$ is shown.
Two plots are displayed: the ratio of the $|t|^2 + |u|^2$ and $|s|^2 + |t|^2 + |u|^2$ approximations over the full calculation, respectively.
In the first case, the approximation is good within $\pm10\%$ over the whole range apart in the low invariant-mass region at both low and large rapidity difference.
The low rapidity difference region possesses remnants of the tri-bosons contributions.
It is therefore expected that the $|t|^2 + |u|^2$ approximations fails in this region.
The second plot, where the $|s|^2 + |t|^2 + |u|^2$ approximation is considered, displays a better behaviour in the previously mentioned region.
The full calculation is approximated at the level of $\pm10\%$ everywhere apart in the upper left corner where the inclusion of $s$-channel contributions did not improve the approximation.
This is due to ...
